---
title: "network visualiser"
author: "christopher barkhuizen"
date: "13/07/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(ggraph)
library(shiny)
library(tidygraph)
library(tidyverse)
library(igraph)
set.seed(12345)
```

## Network Visualiser  

This shiny app generates a network graph based on various input parameters.

The app:   

* Generates a random set of edges      
* Calculates the weight of each edge (based on the count of edges with the same source/destination)   
* Creates a tidy graph object from these edges    
* Plots the graph using ggplot with various graphical parameters based on user input   


## User Input   

Check boxes, selectors and textbox inputs allow the user to:      

* Set the number of vertices
* Set the number of edges 
* Show/hide vertices
* Show/hide vertex labels
* Scale edge width according to the edge weight (number of edges between 2 vertices)
* Change the type of edge (arcs or straight lines)
* Change the colour of edges/vertices/labels

## Sample Network Graph 

```{r tidy_graph}

vertex_count <- 10
edge_count <- 100
edges <- data.frame(src = sample(x = LETTERS[1 : vertex_count], size = edge_count, replace = TRUE), 
                            dest = sample(x = LETTERS[1 : vertex_count], size = edge_count, replace = TRUE))
edges <- group_by(edges, src,dest) %>% summarise(weight = n()) %>% ungroup
nodes <- data.frame(num = 1 : vertex_count, label = LETTERS[1 : vertex_count])
edges <- edges %>% left_join(nodes, by = c("src" = "label")) %>%  
rename(src_id = num) %>% 
left_join(nodes, by = c("dest" = "label")) %>% 
rename(dest_id = num) %>% select(src_id, dest_id, weight)
g_tidy <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE)

g <-   ggraph(g_tidy, layout = "graphopt") + theme_graph()  + 
        geom_edge_link(aes(width = weight), alpha = 0.6, colour = "darkgrey") + 
        scale_edge_width(range = c(0.2, 2.5)) + 
        
        geom_node_point(colour = "black", cex = 8) + 
        geom_node_point(colour = "orange", cex = 7.5) +
        geom_node_text(aes(label = label), colour = "blue", repel = FALSE, size = 5) 
g

cat("shortest path from A (1) to F (6): ",get.shortest.paths(graph = g_tidy, from = 1, to = 6)$vpath[[1]])
```        


## Dependencies and Additional Reading

The app uses: ggplot2, ggraph, shiny, tidygraph, tidyverse  

For more information about network graph visualisations and tidygraph check out:

* [Intro to tidygraph](https://www.data-imaginist.com/2017/introducing-tidygraph/)   
* [Network analysis with R](https://www.jessesadler.com/post/network-analysis-with-r/)   
* [Static and dynamic network visualization with R](http://kateto.net/network-visualization)   



